Insert your title here

### 2018-06-26 mbt.tex
Initial commit in the repository. Created a .gitignore file for not tracking latex output files.

% comment
% This is a general template file for the LaTeX package SVJour3mm
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\usepackage[T1]{fontenc}
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{tkz-graph}
\usepackage{amsmath, amssymb}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{amsfonts}
\usepackage{subcaption}
\usepackage{hyperref}
%\usepackage{pdflscape} % Landscape pages
\usepackage{color}	% Different font colors
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{color}
\usepackage[linesnumbered]{algorithm2e}
\SetKwRepeat{Do}{do}{while}

@@

\section{Introduction}
\label{intro}
The minimum broadcast time (MBT) problem consists of a set of communication nodes with a subset of source nodes. 
The task is to disseminate a signal to every node in a shortest possible time (delay), while abiding by communication rules.
An \emph{informed} node is a node that has received the signal.
Otherwise, a node is \emph{uninformed}.
At the beginning, the set of informed nodes is exactly the set of sources.
An informed node can send the signal to an uninformed node if the two nodes are located within a communication vicinity of each other.

The time is divided into a finite number of time steps.
Every informed node can, at each time step, forward the signal to at most one uninformed neighbor node.
Therefore, the number of informed nodes can at most be doubled from one step to the next.
%\subsection{Motivation and Related Work}
%This communication protocol differs from various wireless communication models where a signal can be relayed to all nodes within a visibility range of a sender.
This communication protocol appears in various practical application such as communication among computer processors or telephone networks.
%The applications are however not confined to wired networks.
Situations where the signals have to cover large distances typically assume sending the signal to one neighbor at a time.
This is common in satellite communication.

{Literature overview}

Deciding whether an instance of MBT has a solution with delay at most $t$ has been shown to be NP-complete \cite{slater81}. 
For bipartite planar graphs with maximum degree 3, NP-completeness persists even if $t=2$ or there is only one source \cite{jansen95}.
When $t=2$, the problem also remains NP-complete for cubic planar graphs \cite{middendorf93}, grid graphs with maximum degree 3,
complete grid graphs, chordal graphs, and for split graphs \cite{jansen95}. 
The single-source variant of the decision version of MBT is NP-complete for grid graphs with maximum degree 4, and for chordal graphs \cite{jansen95}.
The problem is known to be polynomial in trees \cite{slater81}.
Whether the problem is polynomial or NP-complete for split graphs with a single source was stated as an open questions in \cite{jansen95}, and has to the best of our knowledge not been answered yet.

A number of inexact methods, for both general and special graph classes, have been proposed in the literature during the last three decades.
One of the first works of this category \cite{scheuermann84} 
introduces an exact dynamic programming algorithm based on generating all maximum matchings in an induced bipartite graph.
Additional contributions of \cite{scheuermann84} are heuristic approaches for near optimal broadcasting.
From more recent works we mention \cite{hasson04}, which describes a meta heuristic algorithm for MBT, and provides a comparison with other existing methods.
The communication model is considered in an existing satellite navigation system in \cite{chu17}, where a greedy inexact methods is proposed together with a non-linear mathematical model.
Examples of additional efficient heuristics can be found e.g. in \cite{harutyunyan06,harutyunyan14,wang10,jimborean13}.

Approximation algorithms for MBT are studied in \cite{kortsarz95}. 
The authors argue that methods presented in \cite{scheuermann84} provide no guarantee on the performance, and show that the $n$-node wheel is an example of an unfavourable instance.
They introduce an $\mathcal{O}(\sqrt{n})$-additive approximation algorithm for broadcasting in general graphs.
They further provide approximation algorithms for several graph classes with small separators with approximation ratio proportional to the separator size times $\log n$.
Another algorithm with $\mathcal{O}\left(\frac{\log n}{\log \log n}\right)$-approximation\footnote{All logarithms in this paper are of base 2} ratio is given in~\cite{elkin03}.
Most of the works cited above consider a single source.

A related problem extensively studied in the literature is the minimum broadcast graph problem \cite{grigni91,mcgarvey16}. 
A broadcast graph supports a broadcast from any node to all other nodes in optimal time $\lceil\log n\rceil$.
For a given integer $n$, the problem is to find a broadcast graph of $n$ nodes such that e.g.\ the number of edges, or the maximum node degree, in the graph is minimized.
The authors of \cite{mcgarvey16} study ILP models for $c$-broadcast graphs, a generalization of the problem that allows transmission of the signal to at most $c$ neighbours in a single time step.

\section{Network Model and Definitions}

The communication network is represented by a connected graph $G=(V,E)$ and a subset $S\subseteq V$ referred to as the set of sources, where $|V|=n$. 
% Let $V\left[F\right]=\{v\in V:\text{ some edge in } F \text{ is incident to } v\}, F\subseteq E$.
% For disjoint node sets $V_1,V_S\subseteq V$, we denote by $G\left[V_1,V_2\right]$ the bipartite graph $(V_1\cup V_2,E')$, where  $E'=\{\{u,v\}\in E: u\in V_1,v\in V_2\}$.

\begin{definition} \label{def:broadcasttime}
The \emph{broadcast time} $\tau(G,S)$ of a node set $S\subseteq V$ in $G$ is defined as the smallest integer $t\geq 0$ for which there exist
a sequence $V_0\subseteq\dots\subseteq V_t$ of node sets and a function $\pi:V\setminus S\to V$, such that:
\begin{enumerate}
  \item $V_0=S$ and $V_t=V$, \label{def:boundary}
  \item for all $v\in V\setminus S$, $\{v,\pi(v)\}\in E$, \label{def:edge}
  \item for all $i=1,\ldots,t$ and all $v\in V_i$, $\pi(v)\in V_{i-1}$, and \label{def:parent}
  \item for all $u,v\in V_i\setminus V_{i-1}$, $\pi(u)=\pi(v)$ only if $u=v$. \label{def:unique}
\end{enumerate}
% % \begin{equation*}
% \tau(G,S)=
% \begin{cases}
% 	0, S=V,\\
% 	1+\min\{\tau(G,S\cup V\left[M\right]):M \text{ is a matching in } G\left[S, V\setminus S\right]\}, S\neq V.
% \end{cases}
% \label{eq:btime}
% \end{equation*}
\end{definition}
% %Broadcasting is defined as a sequence of sets $(S=V_0\subseteq\dots\subseteq V_k = V)$ where each $V_i$ represents the nodes informed after time step $i$, $0\leq i\leq t$.
% %For each node $v\in V_i\setminus V_{i-1}$, there exists a single node $\pi(v)\in V_{i-1}$ adjacent to $v$, which forwarded the signal to $v$.
% %Also, for every two nodes $u,v\in V_i\setminus V_{i-1}$ we have $\pi(u)=\pi(v)$.
% %The value $k$ is referred to as \emph{delay}.

Referring to Section \ref{intro}, the node set $V_i$ is the set of nodes that are informed in time step $i$.
Initially, only the sources are informed ($V_0=S$), whereas all nodes are informed after $t$ time steps ($V_t=V$),
and the set of informed nodes is monotonously non-decreasing ($V_{i_1}\subseteq V_i$ for $i=1,\ldots,t$).
The parent function $\pi$ maps each node to the node from which it received the signal.
Conditions \ref{def:edge}--\ref{def:parent} of Definition \ref{def:broadcasttime} thus reflect that the sender is a neighbor node in $G$,
and that it is informed at an earlier time step than the recipient node.
Because each node can send to at most one neighbor node in each time step,
condition \ref{def:unique} states that $\pi$ maps the set of nodes which are informed for the first time in step $i$ to distinct parent nodes.

The optimization problem in question is formulated as follows:
\begin{problem}[\textsc{Minimum Broadcast Time}]\label{prob:min}
Given $G=(V,E)$ and $S\subseteq V$, find $\tau(G,S)$.
% a sequence $(S=V_0\subseteq\dots\subseteq V_{\tau(G,S)}=V)$ and a function $\pi:V\setminus S\to V$, such that for each $v\in V\setminus S:\{v,\pi(v)\}\in E$, and for each $u,v\in V_i\setminus V_{i-1}: \pi(u)=\pi(v)\Leftrightarrow u=v$.
\end{problem}

\begin{definition} \label{def:broadcastgraph}
For any $V_0,\ldots,V_t$ and $\pi$ satisfying the conditions of Definition \ref{def:broadcasttime},
the corresponding \emph{communication forest} is
the digraph $D=(V,A)$, where $A=\left\{\left(v,\pi(v)\right): v\in V\right\}$.
Each connected component of $D$ is a \emph{communication tree}.
\end{definition}

\noindent
It is easily verified that the communication trees are indeed arborescences, rooted at distinct sources, with arcs pointing away from the source.
Let $T_s=\left(V_{T_s},A_{V_s}\right)$ denote the communication tree rooted at source $s\in S$, and define
the function $\alpha:V\to S$ such that $v\in V_{T_{\alpha(v)}}$ for all $v\in V$.
That is, $\alpha$ associates a node $v$ with the source in which the communication tree containing $v$ is rooted.
For any communication tree $T_s$, let $T^i_s$ be the subtree of $T$ induced by the set of nodes with distance at most $i$ edges from $s$.
Analogously, let $D^i=\left\{T_s^i: s\in S\right\}$.

% Given a feasible sequence of node sets and $\pi$, we derive the following broadcast trees \cite{grigni91}:
% For $s\in S$, a broadcast tree $T_s=(V_{T_s},E_{T_s})$ is a time-labeled directed subgraph of $G$ describing a broadcast originated in $s$ by the rules:
% \begin{enumerate}
% \item $T_s$ is rooted at $s$ with arcs directed towards the leaves.
% \item each node $v\in V_i$ is labeled with an integer $\tau(v)=i$,
% \item for $v\in V\setminus S, v\in V_{T_s}\Leftrightarrow \pi(v)\in V_{T_s}$, 
% \item $E_{T_s}=\{\{u,v\}: u,v\in V_{T_s}, \pi(v)=u\}$, and
% \item for two sources $s,s'\in S, s\neq s'$ we have that $V_{T_s}\cap V_{T_{s'}}=\emptyset$. 
% \end{enumerate}

% Any feasible sequence of node sets and $\pi$ yield a set of trees $T=\{T_s:s\in S\}$ that forms a partition of $G$ into trees referred to as \emph{broadcast forest}.

% Given a broadcast forest $T$, the delay is determined as $\max_{v\in V}\{\tau(v)\}$.
%A set of trees $T=\{T_s:\cup_{s\in S}V_{T_s}=V\}$ forms a partition of $G$ into trees referred to as \emph{broadcast forest}.
%$T$ can be derived from a given sequence of sets of nodes defining broadcasting and the mapping $\pi$.
% For any integer $i$, let $T^i_s$ be the subtree of $T_s$ obtained by pruning all nodes $v\in V_{T_s}$ with $\tau(v)>i$.
% Analogously, we define $T^i=\{T^i_s:s\in S\}$.

%We also define the set $A=\{(i,j),(j,i):\{i,j\}\in E\}$ that consists of all arcs that can be derived by directing edges in $E$.
The degree of node $v$ in graph $G$ is denoted by $\text{deg}_G(v)$.
For the diameter of $G$ we use the symbol $\Delta_G$.
The set of neighbors of $v\in V$ in $G$ is denoted by $N_G(v)$.
Whenever there is no risk of confusion, the subscript $G$ is omitted.
Likewise, $\text{deg}^-_D(v)$ and $\text{deg}^+_D(v)$ denote, respectively, the in-degree and the out-degree of digraph $D$,
and $N_D^+(v)=\left\{u\in V:(v,u)\in A\right\}$ and $N_D^-(v)=\left\{u\in V:(u,v)\in A\right\}$ denote the neighbor sets of node $v$ in $D$.
The node set $L(D)=\left\{v\in V: N_D^+(v)=\emptyset\right\}$ is referred to as the \emph{leaves} in $D$.

%For convenience, we consider the following definition of broadcast trees \cite{grigni91}:
%For $s\in S$, a broadcast tree $T_s$ with node set $V_{T_s}$ is a time-labeled directed subgraph of $G$ describing a broadcast originated by $s$ by the following rules:
%\begin{enumerate}
%\item $T_s$ is rooted at $s$ with arcs directed towards the leaves.
%\item Each node $v$ is labeled with an integer $\tau(v)$, where $\tau(s)=0$.
%\item Whenever $v$ is a parent of $u$ in $T_s$ $\tau(v)<\tau(u)$.
%\item Whenever $v$ and $u$ are siblings in $T_s$, $\tau(v)\neq \tau(u)$. 
%\end{enumerate}
%The number of leaves in $T_s$ is denoted by $L(T_s)$.
\input{sect3-exact-methods.tex}

\section{Lower bounds}
In this section, we study lower bounds on the delay for several restrictions of input graphs.
%An optimal solution is obtained by solving a sequence of decision problems with varying $t$. 
%It is therefore desirable to determine tight lower and upper bounds in order to arrive in the optimum after solving as few decision problems as possible.
Obvious bounds for a general graph instance are given by
\begin{observation}
For an instance $(G,S)$ of Problem \ref{prob:min},
\begin{equation}
\left\lceil\log\frac{n}{|S|}\right\rceil\leq t^* \leq n-|S|.
\label{eq:loglb}
\end{equation}
\label{obs:loglb}
\end{observation}

@@

\begin{proposition}
For a $d$-regular graph on $n$ nodes and $|S|$ sources, a lower bound on the delay is 
\begin{equation*}
\label{lem:lbreg1}
\underline{t}=\left\lceil\min\{k:2s\sum\limits_{j=1}^k F^{(d-1)}_j\geq n\}\right\rceil.
\end{equation*}
\label{prop:lbfib}
\end{proposition}
\begin{proof}
In order to inform $n$ nodes in the best possible scenario, the signal has to be relayed sufficiently many time steps so that the maximum possible number of informed nodes becomes $n$.
The maximum number of nodes informed within $i$ time steps is given by \eqref{eq:fibcnt}.
We therefore need to set the upper limit of the summation in \eqref{eq:fibcnt} so that the right-hand side exceeds $n$.
The reason why the result is divided by $|S|$ is that the best case scenario with several source nodes assumes that the signals initiated in individual sources are spread evenly.
\qed
\end{proof}

An additional knowledge of a degree sequence of $G$ can be exploited. 
\begin{lemma}
\label{lemma:degorder}
Let $(d_1,\dots,d_n)$ be a degree sequence and $t\leq \bar{t}-|S|$ a given delay.
Consider all possible instances of which graphs are described by this degree sequence.
The maximum number of nodes informed within $t$ time steps is achieved in an instance in which it is possible to inform  nodes in the order of their decreasing degree.

In other words, the instance with the shortest possible broadcast time is such that its nodes can be informed in the order of their decreasing degree.

\end{lemma}
\begin{proof}
A node with degree $d_i$ informed in $i$-th time step can inform at most $\min\{d_i-1,t-i\}$ nodes.
Consider $l,k$ such that $1\leq k < \ell\leq n$ and nodes $v_k$, $v_\ell$ with $\deg(v_k)=d_k$ and $\deg(v_\ell)=d_\ell$.
Then, $d_k-1>d_\ell-1$ and $t-k > t-\ell$.
If nodes are informed in the order of their decreasing degree, let $K_1$ and $L_1$ be the number of nodes that can be informed by $v_k$ and $v_l$, respectively:
$$
K_1=\min\{d_k-1,t-k\}, ~~~ L_1=\min\{d_\ell-1,t-\ell\}.
$$
If the order in which $v_k$ and $v_l$ are informed is switched, i.e. when $v_k$ and $v_\ell$ is informed in the $\ell$-th and $k$-th time step, respectively, 
we use $K_2$ and $L_2$ to denote the maximum number of nodes informed by $v_k$ and $v_\ell$:
$$
K_2=\min\{d_k-1,t-\ell\} ~~~ L_2=\min\{d_\ell-1,t-k\}.
$$
We now investigate what possible values can the expression 
\begin{equation}
\label{eq:degbounds}
(K_1+L_1)-(K_2+L_2)
\end{equation}
attain.
Observe that $K_1\geq L_2$ and $K_2\geq L_1$. 
If $K_2>L_1$, then $K_2=K_1=d_k-1$.
But in that case, $L_2=L_1=d_\ell-1$. 
We conclude that \eqref{eq:degbounds} always takes a non-negative value.
Therefore, if the nodes are informed in the order of their decreasing degrees,
the number of informed nodes within the give time step is no worse than when the nodes are informed in any other order.
\qed
\end{proof}

@@

\begin{algorithm}
\KwData{$n,m,d_1,\dots,d_n\in \mathbb{N}, m\leq n,\newline 1\leq d_n\leq\dots\leq d_2\leq d_1$}
$a_1\leftarrow\dots\leftarrow a_{m}\leftarrow 1$\;%\tcp{this is a comment}
$a_{m+1}\leftarrow\dots\leftarrow a_{n}\leftarrow 0$\;%\tcp{this is a comment}
$c\leftarrow m$;%\tcp{every source informs a new node}
$~k\leftarrow 0$\;
\While {$c<n$} {
$k\leftarrow k+1$\;
$c'\leftarrow 0$\;
\For{$i=1,\dots,c$} {
	\If {$a_i<d_i$} {
		$a_i\leftarrow a_i + 1$\;
		$c'\leftarrow c' + 1$\;
		\If {$c+c'\leq n$} {
			$a_{c+c'}\leftarrow 1$;\tcp{Newly informed node} 
		}
	}
}
$c\leftarrow c + c'$\;
}
\Return $k$\;
%\Return $\lceil k/s \rceil$\;
 \caption{Lower bound exploiting distribution of degrees}
\label{alg:dreg}
\end{algorithm}

The input is assumed to be correct in the sense that a graph with the given degree sequence exists, and by definition, the degree sequence is ordered non-increasingly.
For each iteration $k$, Alg.~\ref{alg:dreg} stores degrees of nodes in $T^k$ in variables $a_1,\dots,a_n$.
Note that the forest $T^k$ is not actually constructed. 
The algorithm operates merely with potential degrees of nodes in $T^k$.
Next, variable $c$ keeps the value $|V_k|$, i.e., the number nodes informed within $k$ steps.
Finally, $c'$ stores $|V_k\setminus V_{k-1}|$, thus the number of nodes newly informed in time step $k$.

>>>>

% This is a general template file for the LaTeX package SVJour3mm
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{tkz-graph}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{amsfonts}
\usepackage{subcaption}
\usepackage{hyperref}
%\usepackage{pdflscape} % Landscape pages
\usepackage{color}	% Different font colors
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{color}
\usepackage[linesnumbered]{algorithm2e}
\SetKwRepeat{Do}{do}{while}

@@

\section{Introduction}
\label{intro}
The minimum broadcast time (MBT) problem consists of a set of communication nodes with a subset of source nodes. 
The task is to disseminate a signal to every node in a shortest possible time $t^*$ (delay), while abiding by communication rules.
An \emph{informed} node is a node that has received the signal.
Otherwise, a node is \emph{uninformed}.
At the beginning, the set of informed nodes is exactly the set of sources.
An informed node $u$ can send the signal to an uninformed node $v$ if $u$ and $v$ are located within a communication vicinity of each other.

The time is divided into discrete time steps.
At each time step, every informed node can forward the signal to at most one uninformed neighbor.
The number of informed nodes at some time step can therefore be no more than the double of the number of informed nodes at the previous time step.
%\subsection{Motivation and Related Work}
%This communication protocol differs from various wireless communication models where a signal can be relayed to all nodes within a visibility range of a sender.
This communication protocol appears in various practical application such as communication among computer processors or telephone networks.
%The applications are however not confined to wired networks.
Situations where the signals have to cover large distances typically assume sending the signal to one neighbor at a time.
This is common in satellite communication.

\subsection{Literature overview}

MBT has been shown to be NP-complete on arbitrary graphs in \cite{slater81}. 
The problem remains NP-complete even for several restricted communication networks \cite{jansen95} such as
split graphs with delay $t^*=2$, chordal graphs,
bipartite and planar graphs with maximum degree 3 and a single source,
several variants of grid graphs,
bipartite planar graphs with delay $t^*=2$ and maximum degree 3. 
Yet another NP-completeness result for 3-regular planar graphs with delay 2 is shown in \cite{middendorf93}.
The problem is known to be polynomial in trees \cite{slater81}.
Whether the problem is polynomial or NP-complete for split graphs with a single source was stated as an open questions in \cite{jansen95}, and hasn't been answered yet to our best knowledge.

A number of inexact methods for both general and special graph classes has been proposed in the literature during the last three decades.
One of the first works of this category is \cite{scheuermann84}, 
where the authors introduce an exact dynamic programming algorithm based on generating all maximum matchings in an induced bipartite graph.
Additional contribution of \cite{scheuermann84} are heuristic approaches for near optimal broadcasting.
From more recent works we mention \cite{hasson04}, which describes a meta heuristic algorithm for MBT, and provides a comparison with other existing methods.
The communication model is considered in an existing satellite navigation system in \cite{chu17}, where a greedy inexact methods is proposed together with a non-linear mathematical model.
Examples of additional efficient heuristics can be found e.g. in \cite{harutyunyan06,harutyunyan14,wang10,jimborean13}.

Approximation algorithms for MBT are studied in \cite{kortsarz95}. 
The authors argue that methods presented in \cite{scheuermann84} provide no guarantee on the performance and state an $n$-node wheel as an example of an unfavourable instance.
They introduce an $\mathcal{O}(\sqrt{n})$-additive approximation algorithm for broadcasting in general graphs.
They further provide approximation algorithms for several graph classes with small separators with approximation ratio proportional to the separator size times $\log n$.
Another algorithm with $\mathcal{O}\left(\frac{\log n}{\log \log n}\right)$-approximation\footnote{All logarithms in this paper are of base 2} ratio is given in~\cite{elkin03}.
Most of the works above consider a single source.

A related problem extensively studied in the literature is the minimum broadcast graph \cite{grigni91,mcgarvey16}. 
A broadcast graph supports a broadcast from any node to all other nodes in optimal time $\lceil\log n\rceil$.
For a given $n$, the effort is to determine properties such as the minimum number of edges and the minimum degree in broadcast graph on $n$ nodes.
The authors of \cite{mcgarvey16} study ILP models for $c$-broadcast graphs, a generalization of the problem that allows transmission of the signal to at most $c$ neighbours in a single time step.

\section{Network Model and Notation}

The communication network is represented by a connected graph $G=(V,E)$ and a subset $S\subseteq V$ referred to as the set of sources, where $|V|=n$. 
Let $V\left[F\right]=\{v\in V:\text{ some edge in } F \text{ is incident to } v\}, F\subseteq E$.
For disjoint node sets $V_1,V_S\subseteq V$, we denote by $G\left[V_1,V_2\right]$ the bipartite graph $(V_1\cup V_2,E')$, where  $E'=\{\{u,v\}\in E: u\in V_1,v\in V_2\}$.

\begin{definition}
The \emph{broadcast time} $\tau(G,S)$ of a node set $S\subseteq V$ in $G$ is defined as: 
\begin{equation*}
\tau(G,S)=
\begin{cases}
	0, S=V,\\
	1+\min\{\tau(G,S\cup V\left[M\right]):M \text{ is a matching in } G\left[S, V\setminus S\right]\}, S\neq V.
\end{cases}
\label{eq:btime}
\end{equation*}
\end{definition}
%Broadcasting is defined as a sequence of sets $(S=V_0\subseteq\dots\subseteq V_k = V)$ where each $V_i$ represents the nodes informed after time step $i$, $0\leq i\leq t$.
%For each node $v\in V_i\setminus V_{i-1}$, there exists a single node $\pi(v)\in V_{i-1}$ adjacent to $v$, which forwarded the signal to $v$.
%Also, for every two nodes $u,v\in V_i\setminus V_{i-1}$ we have $\pi(u)=\pi(v)$.
%The value $k$ is referred to as \emph{delay}.
The optimization problem in question is formulated as follows \cite{jansen95,middendorf93}:
\begin{problem}\label{prob:min}
Given $G=(V,E)$ and $S\subseteq V$, find a sequence $(S=V_0\subseteq\dots\subseteq V_{\tau(G,S)}=V)$ and a function $\pi:V\setminus S\to V$, such that for each $v\in V\setminus S:\{v,\pi(v)\}\in E$, and for each $u,v\in V_i\setminus V_{i-1}: \pi(u)=\pi(v)\Leftrightarrow u=v$.
\end{problem}

Given a feasible sequence of node sets and $\pi$, we derive the following broadcast trees \cite{grigni91}:
For $s\in S$, a broadcast tree $T_s=(V_{T_s},E_{T_s})$ is a time-labeled directed subgraph of $G$ describing a broadcast originated in $s$ by the rules:
\begin{enumerate}
\item $T_s$ is rooted at $s$ with arcs directed towards the leaves.
\item each node $v\in V_i$ is labeled with an integer $\tau(v)=i$,
\item for $v\in V\setminus S, v\in V_{T_s}\Leftrightarrow \pi(v)\in V_{T_s}$, 
\item $E_{T_s}=\{\{u,v\}: u,v\in V_{T_s}, \pi(v)=u\}$, and
\item for two sources $s,s'\in S, s\neq s'$ we have that $V_{T_s}\cap V_{T_{s'}}=\emptyset$. 
\end{enumerate}

Any feasible sequence of node sets and $\pi$ yield a set of trees $T=\{T_s:s\in S\}$ that forms a partition of $G$ into trees referred to as \emph{broadcast forest}.
The number of leaves leaves in $T_s$ is denoted by $L(T_s)$.
The function $\alpha:V\to S$ associates a node $v$ with the source in which the tree containing $v$ is rooted.

Given a broadcast forest $T$, the delay is determined as $\max_{v\in V}\{\tau(v)\}$.
%A set of trees $T=\{T_s:\cup_{s\in S}V_{T_s}=V\}$ forms a partition of $G$ into trees referred to as \emph{broadcast forest}.
%$T$ can be derived from a given sequence of sets of nodes defining broadcasting and the mapping $\pi$.
For any integer $i$, let $T^i_s$ be the subtree of $T_s$ obtained by pruning all nodes $v\in V_{T_s}$ with $\tau(v)>i$.
Analogously, we define $T^i=\{T^i_s:s\in S\}$.

%We also define the set $A=\{(i,j),(j,i):\{i,j\}\in E\}$ that consists of all arcs that can be derived by directing edges in $E$.
The degree of node $v$ in $G$ is denoted by $\text{deg}_G(v)$.
If $G$ is directed, $\text{deg}^-_G(v)$ and $\text{deg}^+_G(v)$ denote in-degree and out-degree, respectively.
For the diameter of $G$ we use the symbol $\Delta_G$.
Whenever there is no danger of confusion, the subscript $G$ is omitted.
The set of neighbors of $v\in V$ in $G$ is denoted by $N(v)$ and $N^+(v)$ in directed graphs.

%For convenience, we consider the following definition of broadcast trees \cite{grigni91}:
%For $s\in S$, a broadcast tree $T_s$ with node set $V_{T_s}$ is a time-labeled directed subgraph of $G$ describing a broadcast originated by $s$ by the following rules:
%\begin{enumerate}
%\item $T_s$ is rooted at $s$ with arcs directed towards the leaves.
%\item Each node $v$ is labeled with an integer $\tau(v)$, where $\tau(s)=0$.
%\item Whenever $v$ is a parent of $u$ in $T_s$ $\tau(v)<\tau(u)$.
%\item Whenever $v$ and $u$ are siblings in $T_s$, $\tau(v)\neq \tau(u)$. 
%\end{enumerate}
%The number of leaves in $T_s$ is denoted by $L(T_s)$.
\input{models.tex}

\section{Lower bounds}
In this section, we study lower bounds on the delay for several restrictions of input graphs.
%An optimal solution is obtained by solving a sequence of decision problems with varying $t$. 
%It is therefore desirable to determine tight lower and upper bounds in order to arrive in the optimum after solving as few decision problems as possible.
Obvious bounds for a general graph instance are given by
\begin{observation}
For an instance $(G,S)$ of Problem \ref{prob:min},
$$\left\lceil\log\frac{n}{|S|}\right\rceil\leq t^* \leq n-|S|.$$
\end{observation}

@@

\begin{proposition}
For a $d$-regular graph on $n$ nodes and $|S|$ sources, a lower bound on the delay is 
\begin{equation*}
\label{lem:lbreg1}
\underline{t}=\left\lceil\min\{k:2\sum\limits_{j=1}^k F^{(d-1)}_j\geq n\}/|S|\right\rceil.
\end{equation*}
\end{proposition}
\begin{proof}
In order to inform $n$ nodes in the best possible scenario, the signal has to be relayed sufficiently many time steps so that the maximum possible number of informed nodes becomes $n$.
The maximum number of nodes informed within $i$ time steps is given by \eqref{eq:fibcnt}.
We therefore need to set the upper limit of the summation in \eqref{eq:fibcnt} so that the right-hand side exceeds $n$.
The reason why the result is divided by $|S|$ is that the best case scenario with several source nodes assumes that the signals initiated in individual sources are spread evenly.
\qed
\end{proof}

An additional knowledge of a degree sequence of $G$ can be exploited. 
\begin{lemma}
\label{lemma:degorder}
Let $(d_1,\dots,d_n)$ be a degree sequence and $t\leq \bar{t}-|S|$ a given delay.
The maximum number of nodes informed within $t$ time steps is achieved if nodes become informed in the order of their decreasing degree.
\end{lemma}
\begin{proof}
A node with degree $d_i$ informed in $i$-th time step can inform at most $\min\{d_i-1,t-i\}$ nodes.
Consider $l,k$ such that $1\leq k < \ell\leq n$ and nodes $v_k$, $v_\ell$ with $\deg(v_k)=d_k$ and $\deg(v_\ell)=d_\ell$.
Then, $d_k-1>d_\ell-1$ and $t-k > t-\ell$.
If nodes are informed in the order of their decreasing degree, let $K_1$ and $L_1$ be the number of nodes that can be informed by $v_k$ and $v_l$, respectively:
$$
K_1=\min\{d_k-1,t-k\}, ~~~ L_1=\min\{d_\ell-1,t-\ell\}.
$$
If the order in which $v_k$ and $v_l$ are informed is switched, i.e. when $v_k$ and $v_\ell$ is informed in the $\ell$-th and $k$-th time step, respectively, 
we use $K_2$ and $L_2$ to denote the maximum number of nodes informed by $v_k$ and $v_\ell$:
$$
K_2=\min\{d_k-1,t-\ell\} ~~~ L_2=\min\{d_\ell-1,t-k\}.
$$
We now investigate what possible values can the expression 
\begin{equation}
\label{eq:degbounds}
(K_1+L_1)-(K_2+L_2)
\end{equation}
attain.
Observe that $K_1\geq L_2$ and $K_2\geq L_1$. 
If $K_2>L_1$, then $K_2=K_1=d_k-1$.
But in that case, $L_2=L_1=d_\ell-1$. 
We conclude that \eqref{eq:degbounds} always takes a non-negative value.
Therefore, if the nodes are informed in the order of their decreasing degrees,
the number of informed nodes within the give time step is no worse than when the nodes are informed in any other order.
\qed
\end{proof}

@@

\begin{algorithm}
\KwData{$n,m,d_1,\dots,d_n\in \mathbb{N}, m\leq n,\newline 1\leq d_n\leq\dots\leq d_2\leq d_1$}
$a_1\leftarrow\dots\leftarrow a_{m}\leftarrow 1$\;%\tcp{this is a comment}
$a_{m+1}\leftarrow\dots\leftarrow a_{n}\leftarrow 0$\;%\tcp{this is a comment}
$c\leftarrow m$;%\tcp{every source informs a new node}
$~k\leftarrow 0$\;
\While {$c<n$} {
$k\leftarrow k+1$\;
$c_n\leftarrow 0$\;
\For{$i=1,\dots,c$} {
	\If {$a_i<d_i$} {
		$a_i\leftarrow a_i + 1$\;
		$c_n\leftarrow c_n + 1$\;
		\If {$c+c_n\leq n$} {
			$a_{c+c_n}\leftarrow 1$;\tcp{Newly informed node} 
		}
	}
}
$c\leftarrow c + c_n$\;
}
\Return $k$\;
%\Return $\lceil k/s \rceil$\;
 \caption{Lower bound exploiting distribution of degrees}
\label{alg:dreg}
\end{algorithm}

The input is assumed to be correct in the sense that a graph with the given degree sequence exists, and by definition, the degree sequence is ordered non-increasingly.
For each iteration $k$, Alg.~\ref{alg:dreg} stores degrees of nodes in $T^k$ in variables $a_1,\dots,a_n$.
Note that the forest $T^k$ is not actually constructed. 
The algorithm operates merely with potential degrees of nodes in $T^k$.
Next, variable $c$ keeps the value $|V_k|$, i.e., the number nodes informed within $k$ steps.
Finally, $c_n$ stores $|V_k\setminus V_{k-1}|$, thus the number of nodes newly informed in time step $k$.
